The following are the initial codes given. It mainly purposed to clear existing objects,and create individual data.

```{r}
rm(list = ls()) 
student_id <- 33521352
set.seed(student_id) # Your Student ID is the random seed 
WD = read.csv("WinnData.csv") 
WD = WD[sample(nrow(WD), 5000, replace = FALSE),] 
WD = WD[,c(sort(sample(1:30,20, replace = FALSE)), 31)] 
options(warn = -1)
```

```{r}
# The way of installing packages is adopted from https://stackoverflow.com/questions/4090169/elegant-way-to-check-for-missing-packages-and-install-them

list.of.packages <- c("adabag", "rpart", "e1071", "tree", "ROCR", "smotefamily", "neuralnet", "lightgbm")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

library(dplyr)
library(tree)
library(ipred)
library(rpart)
library(e1071)
library(neuralnet)
library(ROCR)
library(lightgbm)
```

**Question 1**

Investigation of oats propotions:

```{r}
class_prop <- prop.table(table(WD$Class))
```

1.1 - Showing propotions

```{r}
cat("Proportion of Class 1:", class_prop[2] * 100 , "% \n")
cat("Proportion of Class 0:", class_prop[1]* 100 , "% \n")
```

1.2 - Descriptive Stastistics

```{r}
# Since we want descriptions for all INDEPENDENT variables, this means we dont want Class column to be exist in our descriptive statistics. That is,
# 1. We select those non-classes cases, AND THEN
# 2. We do descriptive statistics from descr functions manually

desc <- WD %>%
  select(-Class) %>%
  {
    data.frame(
      NA_Count = sapply(., function(x) sum(is.na(x))),
      Mean     = sapply(., function(x) mean(x)),
      Std_Dev  = sapply(., function(x) sd(x)),
      Q1       = sapply(., function(x) as.numeric(quantile(x, 0.25))),
      Median   = sapply(., function(x) median(x)),
      Q3       = sapply(., function(x) as.numeric(quantile(x, 0.75))),
      Max      = sapply(., function(x) max(x))
    )
  }


# 3. AND THEN, make it dataframe so it's nice looking

desc <- desc %>% 
  round(2)%>%
  t() %>% 
  as.data.frame()


desc
```

**Question 2**

No any pre-processing since there exist **no** missing variables

The double checking below is to ensure that `Class` is a factor

```{r}
WD$Class <- as.factor(WD$Class)  
```

**Question 3**

As instructed by Information Sheet

```{r}
set.seed(student_id) 
train.rows = sample(1:nrow(WD), 0.7*nrow(WD)) 
data.train = WD[train.rows,] 
data.test = WD[-train.rows,] 
```

```{r}
is.factor(data.train$Class)
levels(data.train$Class)  
str(data.train$Class)
```

**Question 4**

Decision Tree

```{r}

tree_model <- tree(Class ~ ., data = data.train)

```

Naïve Bayes

```{r}

nb_model <- naiveBayes(Class ~ ., data = data.train)

```

Bagging

```{r}
bag_model <- bagging(Class ~ ., data = data.train)
```

Boosting - AdaBoost

```{r}
library(adabag)
boost_model <- boosting(Class ~ ., data = data.train)

```

Random Forest

```{r}
library(randomForest)

rf_model <- randomForest(Class ~ ., data = data.train)
```

**Question 5, 6, 7, 8, 9**

```{r}
library(scales)
metrices <- function(pred, prob, data.train, data.test) {
    ### This is a function that evaluate all necessary metrics including
    ### Accuracy, Precision, Recall, F1_Score
    ### The calculation is mainly based on what was taught on Week 6
    ### But the calculation is basd on confusion matrix since it's asked.
    ### Return:
    ### Whole bunch of list that are contains of required information
    
    Confusion_Matrix = table(Predicted = pred, Actual = data.test$Class)
    
    Accuracy = sum(diag(Confusion_Matrix)) / sum(Confusion_Matrix)
    Precision = Confusion_Matrix[2,2] / sum(Confusion_Matrix[2,])
    Recall = Confusion_Matrix[2,2] / sum(Confusion_Matrix[,2])
    F1_Score = 2 * (Precision * Recall) / (Precision + Recall)
    
    pred_obj = ROCR::prediction(prob[,2], data.test$Class)
    perf = performance(pred_obj, "tpr", "fpr")
    auc = as.numeric(performance(pred_obj, "auc")@y.values)
    
    return(list(Confusion_Matrix = Confusion_Matrix, 
                Accuracy = Accuracy, 
                Precision = Precision, 
                Recall = Recall, 
                F1_Score = F1_Score,
                pred_obj = pred_obj,
                perf = perf,
                auc = auc
                ))
}

evaluate_model <- function(model, data.train, data.test) {
    ### This is a function that evaluate all necessary metrics including
    ### Accuracy, Precision, Recall, F1_Score of
    ### The calculation is mainly based on what was taught on Week 6
    ### The calculation would be based on the 'metrices' function
    ### Return:
    ### Model with information evaluated
  
  model_prob <- model$prob
  model_info <- metrices(pred = model_pred, prob = model_prob, data.train = data.train, data.test = data.test)
  
  model <- append(model, model_info)
  
  return(model)
}

```

```{r}
model_entry <- function(model, pred, prob, name) {
    ### This is a function that make a model entry (global entry) into whole bunch of models
    ### It regiters all the necessary information
    ### Model with information to be appended into the 'whole_bunch_of_models'
  
  # colour
  model_entry <- function(model, pred, prob, name) {
      ### This is a function that make a model entry (global entry) into whole bunch of models
      ### It regiters all the necessary information
      ### Model with information to be appended into the 'whole_bunch_of_models'
    new_entry <- list(
      model = model,
      pred = pred,
      prob = prob,
      name = name,
      colour = hue_pal()(length(whole_bunch_of_models) + 1)[length(whole_bunch_of_models) + 1]
    )
    entry <- evaluate_model(
      model = new_entry,
      data.train = data.train,
      data.test = data.test
    )
    
    whole_bunch_of_models <<- append(whole_bunch_of_models, list(entry))
}}


### These are whole bunch of list of models
### I'm making it list so that it's extendable
### And cz every model have dofferent type of method to call predict
whole_bunch_of_models <- list()

model_entry(model = tree_model, 
       pred = predict(tree_model, data.test, type = "class"),
       prob = predict(tree_model, data.test, type = "vector"),
       name = "Decision Tree")
  
model_entry(model = nb_model,
       pred = predict(nb_model, data.test),
       prob = predict(nb_model, data.test, type = "raw"),
       name = "Naive Bayes")
  
model_entry(model = bag_model, 
       pred = predict(bag_model, data.test, type= "class"),
       prob = predict(bag_model, data.test, type= "prob"),
       name = "Bagging")
  
model_entry(model = boost_model,
       pred = predict(boost_model, data.test)$class,
       prob = predict(boost_model, data.test)$prob,
       name = "Boosting")
  
model_entry(model = rf_model, 
       pred = predict(rf_model, data.test),
       prob = predict(rf_model, data.test, type = "prob"),
       name = "Random Forest")
```

```{r}
plot_roc <- function(whole_bunch_of_models) {
  # This is a function that plot ROC curve and return ROC curve when needed.
  # It uses all models in 'whole_bunch_of_models'
  plot(NA, xlim = c(0, 1), ylim = c(0, 1), 
       xlab = "False Positive Rate", ylab = "True Positive Rate",
       main = "ROC Curves Comparison")
  
  
  abline(0, 1, col = "grey")
  
  
  for (model in whole_bunch_of_models) {
    plot(model$perf, col = model$colour, add = TRUE)
  }
  
  
  legend("bottomright",
         legend = paste0(sapply(whole_bunch_of_models, function(model) model$name), 
                        " (AUC = ", 
                        sapply(whole_bunch_of_models, function(model) round(model$auc, 3)), ")"),
         fill = sapply(whole_bunch_of_models, function(model) model$colour), 
         cex = 0.8)
}
```

```{r}
compare_metrices <- function (whole_bunch_of_models){
  # This is a function that return a dataframe that compare all models in 'whole_bunch_of_models'
  # It gives all necessary information needed. 
    metrices_list <- list()
  
  for (models in whole_bunch_of_models) {
    metrices_list[[models$name]] <- data.frame(
      Accuracy = models$Accuracy,
      Precision = models$Precision,
      Recall = models$Recall,
      F1_Score = models$F1_Score
    )
  }
  
  metrices_compare <- do.call(rbind, metrices_list)
  metrices_compare <- (metrices_compare  
                       %>% round(2) 
                       %>% replace_na(replace = list(Accuracy = 0, Precision = 0, Recall = 0, F1_Score = 0))%>% t() %>% as.data.frame())
  return (
    metrices_compare
  )
}
```

```{r}
confusion_matrices <- function(whole_bunch_of_models) {
  ### This is a function that return all confusion matrices of all models when required based on models in 'whole_bunch_of_models'
  for (models in whole_bunch_of_models) {
    cat("\n")
    cat("=======================\n")
    cat("Confusion Matrix of", models$name, "\n")
    cat("=======================\n")
    print(models$Confusion_Matrix)
    cat("\n")
  }
}
```

```{r}
confusion_matrices(whole_bunch_of_models)
```

```{r}
library(tidyr)

compare_metrices(whole_bunch_of_models)
```

```{r}
plot_roc(whole_bunch_of_models)
```

**Q9-10**

```{r}
# Use Random Forest since its AUC is largest
importance(rf_model)
varImpPlot(rf_model)
```

```{r}
# Boosting model is also referred
importanceplot(boost_model)
```

```{r}
# This is quite obviously, sorting the model importance in decreasing order
sorted_importance <- sort(importance(rf_model)[, "MeanDecreaseGini"], decreasing = TRUE)
sorted_importance
```

```{r}
# Take top first and top thrid due to the same reason mentioned in the report
top_n <- 3
important_features <- names(sorted_importance)[1:top_n]
thresholds <- sapply(important_features, function(f) {
  median(WD[[f]][WD$Class == 1], na.rm = TRUE) %>% round(1)
})

```

```{r}
hand_model <- function(features, thresholds) {
  list(features = features, thresholds = thresholds)
}
# simplified prediction function
predict.hand_model <- function(model, newdata) {
  f1 <- model$features[1]
  f2 <- model$features[3]
  t1 <- model$thresholds[1]
  t2 <- model$thresholds[3]
  
  # here are the decision rule based on defined rules mentioned in report
  # pred_class and prob are to make sure that they could be compatible in the 'whole_bunch_of_models' list.
  pred_class <- ifelse(newdata[[f1]] > t1 | newdata[[f2]] < t2, 1, 0)
  prob <- ifelse(pred_class == 1, 0.99, 0.01)  # Simple confidence values
  
  list(class = factor(pred_class, levels = c(0, 1)),
       prob = data.frame(`0` = 1 - prob, `1` = prob))
}

# evaluate model
manual_model <- hand_model(important_features, thresholds)
preds <- predict.hand_model(manual_model, data.test)

model_entry(model = manual_model,
            pred = preds$class,
            prob = preds$prob,
            name = "Simple Rules")
```

```{r}
# Here is the rules
cat("Manual Prediction Rules:\n",
    "1. IF", important_features[1], ">", thresholds[[1]], "THEN predict 'Oats' (Class = 1)\n",
    "2. ELSE IF", important_features[3], "<", thresholds[[3]], "THEN predict 'Oats' (Class = 1)\n",
    "3. OTHERWISE predict 'Other' (Class = 0)\n")

```

```{r}
confusion_matrices(whole_bunch_of_models)
```

```{r}
compare_metrices(whole_bunch_of_models)
```

```{r}
plot_roc(whole_bunch_of_models)
```

**Q11**

```{r}
# 1. Normalise
WD_scaled <- scale(WD[,1:length(names(WD))-1]) %>% as.data.frame()
WD_scaled$Class <- as.factor(WD$Class)
head(WD_scaled)
```

```{r}
library(smotefamily)
smote_result <- SMOTE(
  X = WD_scaled %>% select(-Class),
  target = WD_scaled$Class,
  dup_size = 0)

WD_balanced <- smote_result$data

# Here idk why they renamed "Class" to "class" 
# therefore here is whete im renaming it back
WD_balanced <- WD_balanced %>% rename(Class = class)
WD_balanced$Class <- as.factor(WD_balanced$Class)
```

```{r}
cat("Ori class distribution:\n")
print(table(WD$Class))

cat("\nBalanced class distribution:\n")
print(table(WD_balanced$Class))
```

```{r}
# 2. Again train-test split
# But there exist imbalance of data between Class 0 and Class 1, therefore we are taking stratified samples...
library(caret)

set.seed(student_id) 
train.rows <- sample(1:nrow(WD_balanced), 0.7*nrow(WD_balanced)) 
data.train <- WD_balanced[train.rows, ]
data.test <- WD_balanced[-train.rows, ]
```

```{r}
train_control <- trainControl(method = "cv", 
                              number = 10)

```

```{r}
# 3. AND THEN, do grid-search for yperparameyer tuning
# Since we dk the best number of features per split (mtry)
# and the best node size of the random forest
# we would do the following tuning on
# mtry : 2 to number of features/3 + 1
# nodesize : 2 to 8

mtry_grid <- 2: ceiling(ncol(data.train)/3) + 1
nodesize_grid <- 2: 5

results <- list()

for (i in seq_along(mtry_grid)) {
  for (j in seq_along(nodesize_grid)) {
    mtry = mtry_grid[i]
    nodesize = nodesize_grid[j]
      
    model <- randomForest(
      Class ~ .,
      data = data.train,
      trControl = train_control,
      mtry = mtry,
      nodesize = nodesize,
      ntree = 500  # Fixed ntree,
      
    )
    
    
    pred <- predict(model, data.test)
    
    
    Confusion_Matrix <- table(Predicted = pred, Actual = data.test$Class)
    Accuracy <- sum(diag(Confusion_Matrix)) / sum(Confusion_Matrix)
    
    
    results[[length(results)+1]] <- data.frame(
      mtry = mtry,
      nodesize = nodesize,
      Accuracy = round(Accuracy, 3)
    )
  }
}


results_df <- do.call(rbind, results) %>% arrange(desc(Accuracy))

```

```{r}
head(results_df,5)
```

```{r}
best_params <- results_df[1,]
print("Best mtry paraneter")
best_params$mtry[1]
print("Best nodesize parameter")
best_params$nodesize
```

```{r}
optimised_model <- randomForest(
      Class ~ .,
      data = data.train,
      mtry = best_params$mtry[1],
      nodesize = best_params$nodesize[1],
      trControl = train_control,
      ntree = 500
    )

```

```{r}
optimised_model
```

```{r}
model_entry(optimised_model, 
       pred = predict(optimised_model, data.test),
       prob = predict(optimised_model, data.test, type = "prob"),
       name = "Optimised Random Forest")
```

```{r}
compare_metrices(whole_bunch_of_models)
```

```{r}
plot_roc(whole_bunch_of_models)
```

```{r}
optimised_model
```

**Q12**

```{r}
# According to the current insights
# it is proven that SMOTE resampling is necessary.
# therefore, the the ANN will be based on the balanced_data

nn_model <- neuralnet(Class ~., 
                      data = data.train,
                      hidden = 1)

nn_compute <- compute(nn_model,data.test[,1:length(data.test)-1])

nn_pred <- round(nn_compute$net.result, 0)[,2]

nn_prob <- nn_compute$net.result
```

```{r}
model_entry(
  nn_model,
  pred = nn_pred,
  prob = nn_prob,
  name = "ANN"
)
```

```{r}
compare_metrices(whole_bunch_of_models)
```

```{r}
plot_roc(whole_bunch_of_models)
```

**Q13**

```{r}
# Based on https://lightgbm.readthedocs.io/en/latest/R/index.html
train_data <- as.matrix(data.train[, -which(names(data.train) == "Class")])
test_data <- as.matrix(data.test[, -which(names(data.test) == "Class")])

# Create LightGBM dataset objects
dtrain <- lgb.Dataset(data = train_data, label = as.numeric(data.train$Class)-1)
dtest <- lgb.Dataset.create.valid(dtrain, data = test_data, label = as.numeric(data.test$Class)-1)
```

```{r}
lgb_model_baseline <- lgb.train(
  data = dtrain,
  nrounds = 250
)
```

```{r}
preds = predict(lgb_model_baseline, test_data)
model_entry(
  lgb_model_baseline,
  pred = round(preds, 0),
  prob = data.frame('0' = 1-preds, '1' = preds),
  name = "Baseline LightGBM model"
)
```

```{r}
#Similar to optimised_random forest, we will be doing hyperparameter tuning for this model too.
# and cv
# The most important predictors of this model are:
# learning rate and number of leaves
# Therefore the hyperparameter tuning will be conducted by:
# num_leaves : 5 random values between 10 to 100
# learning rate: among 0.01, 0.05 and 0.1 --- which are the most common rates used

# while for CV, still, 10-fold will be conducted

grid_num_leaves =  seq(8, 35, by = 5)
grid_learning_rate = c(0.01, 0.05, 0.1)

results <- list()

for (i in seq_along(grid_num_leaves)) {
  for (j in seq_along(grid_learning_rate)) {
    num_leaves = grid_num_leaves[i]
    learning_rate = grid_learning_rate[j]
      
    cv_model <- lgb.cv(
      
      data = dtrain,
      nrounds = 250,
      nfold = 10,
      params = list(
        objective = "binary",
        num_leaves = num_leaves,
        learning_rate = learning_rate
      ),
      verbose = -1 # so that it dont spam words
      
    )
    
    best_nrounds <- cv_model$best_iter
    
    model <- lgb.train(
      params = list(
        num_leaves = num_leaves,
        learning_rate = learning_rate
      ),
      data = dtrain,
      nrounds = best_nrounds,
      verbose = -1 # so that it dont spam words
    )    
    
    
    pred <- round(predict(model, test_data), 0)
    
    
    Confusion_Matrix <- table(Predicted = pred, Actual = data.test$Class)
    Accuracy <- sum(diag(Confusion_Matrix)) / sum(Confusion_Matrix)
    
    
    results[[length(results)+1]] <- data.frame(
      num_leaves = num_leaves,
      learning_rate = learning_rate,
      nrounds = best_nrounds,
      Accuracy = round(Accuracy, 3)
    )
  }
}


results_df <- do.call(rbind, results) %>% arrange(desc(Accuracy))
best_params <- results_df[1,]
```

```{r}
head(results_df,5)
```

```{r}
lgb_model <- lgb.train(
      params = list(
        num_leaves = best_params$num_leaves,
        learning_rate = best_params$learning_rate
      ),
      data = dtrain,
      nrounds = best_params$nrounds,
      verbose = -1 # so that it dont spam words
    )
```

```{r}
preds = predict(lgb_model, test_data)
model_entry(
  lgb_model,
  pred = round(preds, 0),
  prob = data.frame('0' = 1-preds, '1' = preds),
  name = "Optimised LightGBM model"
)
```

```{r}
compare_metrices(whole_bunch_of_models)
```

```{r}
plot_roc(whole_bunch_of_models)
```
